Precision calculations for cosmic shear

### 2017-02-07 cfhtlens.tex
MK edit text

The Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS) represented a major step forward for the field of weak gravitational lensing, in terms of improved accuracy in data reduction \citep{erben/etal:2013}, the implementation of PSF-Gaussianised matched multi-band photometry \citep{hildebrandt/etal:2012}, cross-correlation clustering analysis between photometric redshift slices to verify tomographic redshift distributions \citep{benjamin/etal:2013}, accurate calibrated shape measurements \citep{miller/etal:2013} and a full suite of informative systematic tests to select a clean data set \citep{heymans/etal:2012}.    Since the public release of this survey in 2013, the community has continued to scrutinise and advance our understanding of CFHTLenS by identifying a number of areas where analyses could improve:
\begin{itemize}
 \item{\citet{choi/etal:2016} identified significant biases in the tomographic photometric redshift distributions using a more effective clustering analysis, in comparison to \citet{benjamin/etal:2013}, by incorporating newly overlapping spectroscopy from the Sloan Digital Sky Survey.  The conclusion of this work was that any re-analysis of CFHTLenS should include systematic error terms to account for bias and scatter, with a prediction that accounting for these biases would be {\it reduce} the recovered amplitude of $\sigma_8$ by $\sim 4$\%. }
\item{ The CFHTLenS tomographic cosmological analysis was then revisited by \citet{joudaki/etal:2016} in order to include a full redshift error analysis based on the results from \citet{choi/etal:2016}.  The impact of correcting for these biases, including their associated errors, served to reduce the overall constraining power of the survey and hence also the tension between CFHTLenS and CMB constraints.}
 \item{\citet{asgari/etal:2016} used the stringent `COSEBI' statistic to identify significant non-lensing `B-mode' distortions when the CFHTLenS data was split into tomographic slices.}
\item{\citet{kuijken/etal:2015} showed that the CFHTLenS shear calibration corrections derived in \citet{miller/etal:2013} were underestimated as a result of an imperfect match between the galaxy populations in the data and image simulations.}
\item{\citet{fenechconti/etal:2016} demonstrated that the CFHTLenS data would have been subject to a weight bias that favours galaxies that are more intrinsically oriented with the point-spread function.  They also showed that the impact of calibration selection biases, that were not considered in \citet{miller/etal:2013}, would have lead to the over-correction of multiplicative shear bias in the CFHTLenS analyses, by a few percent.}
\item{\citet{joudaki/etal:2016} updated the CFHTLenS covariance matrices using larger-box numerical simulations that were less subject to the lack of power on large scales.}
\item{\cite{2012ApJ...761..152T} provided a more accurate non-linear power spectrum correction that that used in the original CFHTLenS analyses and the halo model from \cite{2015MNRAS.454.1958M} allowed for simultaneous modelling of baryonic modifications to the non-linear power spectrum.} 
\end{itemize}
All these advances in our understanding were incorporated and accounted for in the recent KiDS cosmic shear analysis \citep{KiDS-450} which reports a $2.3 \sigma$ tension with Planck.  Efforts are now underway to fully re-analyse CFHTLenS using the advanced KiDS analysis pipeline with revised shape measurements and calibrations for the shear and photometric redshifts.  Until this analysis is complete we note that these known shortcomings with the original CFHTLenS results impact in different ways on the cosmological conclusions that one can draw from the survey.  As CFHTLenS has similar statistical power to current weak lensing surveys, however, it is nevertheless provides a very useful testbed with which to demonstrate the impact of adopting different approximations when constraining cosmological parameters.

>>>>

The Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS) represented a
major step forward for the field of weak gravitational lensing, in terms of
improved accuracy in data reduction \citep{CFHTLenS-data}, the implementation
of PSF-Gaussianised matched multi-band photometry
\citep{CFHTLenS-photoz}, cross-correlation clustering analysis between
photometric redshift slices to verify tomographic redshift distributions
\citep{CFHTLenS-2pt-tomo}, accurate calibrated shape measurements
\citep{CFHTLenS-shapes} and a full suite of informative systematic tests to
select a clean data set \citep{CFHTLenS-sys}. Since the public release
of this survey in 2013, the community has continued to scrutinise and advance
our understanding of CFHTLenS by identifying a number of areas where analyses
could improve:
%
\begin{itemize}
%
 \item{\citet{2016MNRAS.463.3737C} identified significant biases in the tomographic
photometric redshift distributions using a more effective clustering analysis,
in comparison to \citet{CFHTLenS-2pt-tomo}, by incorporating newly overlapping
spectroscopy from the Sloan Digital Sky Survey.  The conclusion of this work
was that any re-analysis of CFHTLenS should include systematic error terms to
account for bias and scatter, with a prediction that accounting for these
biases would be {\it reduce} the recovered amplitude of $\sigma_8$ by $\sim
4$\%. }
%
\item{ The CFHTLenS tomographic cosmological analysis was then revisited by
\citet{joudaki/etal:2016} in order to include a full redshift error analysis
based on the results from \citet{2016MNRAS.463.3737C}.  The impact of
correcting for these biases, including their associated errors, served to
reduce the overall constraining power of the survey and hence also the tension
between CFHTLenS and CMB constraints.}
%
 \item{\citet{2016arXiv160100115A} used the stringent `COSEBI' statistic to
identify significant non-lensing `B-mode' distortions when the CFHTLenS data
was split into tomographic slices.}
%
\item{\citet{2015MNRAS.454.3500K} showed that the CFHTLenS shear calibration
corrections derived in \citet{CFHTLenS-shapes} were underestimated as a result
of an imperfect match between the galaxy populations in the data and image
simulations.}
%
\item{\citet{fenechconti/etal:2016} demonstrated that the CFHTLenS data would
have been subject to a weight bias that favours galaxies that are more
intrinsically oriented with the point-spread function.  They also showed that
the impact of calibration selection biases, that were not considered in
\citet{miller/etal:2013}, would have lead to the over-correction of
multiplicative shear bias in the CFHTLenS analyses, by a few percent.}
%
\item{\citet{joudaki/etal:2016} updated the CFHTLenS covariance matrices using larger-box numerical simulations that were less subject to the lack of power on large scales.}
\item{\cite{2012ApJ...761..152T} provided a more accurate non-linear power
spectrum correction that that used in the original CFHTLenS analyses and the
halo model from \cite{2015MNRAS.454.1958M} allowed for simultaneous modelling
of baryonic modifications to the non-linear power spectrum.} 
%
\end{itemize}
%
All these advances in our understanding were incorporated and accounted for in
the recent KiDS cosmic shear analysis \citep{KiDS-450} which reports a $2.3
\sigma$ tension with Planck.  Efforts are now underway to fully re-analyse
CFHTLenS using the advanced KiDS analysis pipeline with revised shape
measurements and calibrations for the shear and photometric redshifts.  Until
this analysis is complete we note that these known shortcomings with the
original CFHTLenS results impact in different ways on the cosmological
conclusions that one can draw from the survey.  As CFHTLenS has similar
statistical power to current weak lensing surveys, however, it is nevertheless
provides a very useful testbed with which to demonstrate the impact of adopting
different approximations when constraining cosmological parameters.

### 2017-02-07 discussion.tex
MK edit text

\citet{kilbinger/etal:2013} present a detailed comparison of cosmological constraints obtained from a range of different two-point shear statistics including the shear correlation function, $\xi_\pm$, the aperture-mass dispersion, $\langle M_ {\rm ap} \rangle ^2$ \citep{schneider/etal:1998}, and the COSEBIs, $E_n$ \citep{schneider/etal:2010}.  These statistics are linearly related to the power spectrum via integrals of the form,
%
\begin{align}
%\label{eqn:integ}
& E_n =\int_0^{\infty}\frac{\d \ell \ell}{2\pi} W_n(\ell) P(\ell)\;,\\ \nonumber
& \langle M_ {\rm ap} \rangle ^2(\theta)=\int_0^{\infty}\frac{\d \ell \ell}{2\pi} U^2_\theta(\ell) P(\ell)\;,
\end{align}
%
where $W_n(\ell)$ and $U^2_\theta(\ell)$ are defined in \cite{schneider/etal:2010} and \cite{schneider/etal:1998}. The corresponding equations for $\xi_\pm$ are given in equation~\ref{eqn:xiGG}.
Figure~\ref{fig:filters} shows the integrands of these statistics 
for two cases, normalised to their maximum value, where the integrands are of the form $\ell F(\ell) P(\ell)$. For $\xi_\pm$, $P(\ell)$ is equal to the sum of the E and B-mode power spectra, motivating the development of the aperture-mass dispersion statistic, $\langle M_ {\rm ap} \rangle ^2$.  This statistic is, however, a lossy conversion and is biased by small angular separations, where blending of galaxies makes shear measurement challenging \citep{kilbinger/etal:2006}.  The COSEBIs statistic tackles both these shortcomings.  The upper two panels of figure~\ref{fig:filters} show the integrands\footnote{Note that the lower left panel of Fig. 1 in \citet{kitching/etal:2016} shows an integrated form of this function for a maximum angular separation of $100'$. However, in \citet{kilbinger/etal:2013}, the data used in \citet{kitching/etal:2016}, $\theta$ is between $0.8'$ and $350'$.} of $\xi_\pm$ for $\theta=100'$ and $\theta=350'$. The lower middle panel in Figure~\ref{fig:filters} shows the COSEBIs integrands for two angular ranges, $[1',100']$ and $[0.8',350']$, where we only show the integrands for the lowest COSEBIs mode, $E_1$, as the higher modes generally probe larger $\ell$-modes.  Finally, the lowest panel shows the integrands of aperture mass dispersion statistics, for the same two maximum angular ranges.

@@

From Figure~\ref{fig:filters} we can see that the two-point cosmic shear statistics tested by \citet{kilbinger/etal:2013} exhibit different dependences between the angular scales sampled and the $\ell$-range probed.   
If a significant bias had been introduced at low-$\ell$ by using flat-sky and Limber approximations, we would then expect to see a systematic shift between the different two-point statistics with the COSEBIs statistic being essentially unaffected as it only includes modes with $\ell \gtrsim 20$.  This is found not to be the case with all three statistics finding $\sigma_8 (\Omega_m/0.27)^\alpha = 0.79$ with errors that range from $0.03$ to $0.06$ for the full five-parameter fit, and $\alpha$ ranging from $0.59$ to 0.7 \citep[see Table 5 of][]{kilbinger/etal:2013}.  This comparison further supports our argument that the approximations highlighted by \citet{kitching/etal:2016} have negligible impact for current surveys.

>>>>

\citet{CFHTLenS-2pt-notomo} present a detailed comparison of cosmological
constraints obtained from a range of different two-point shear statistics
including the shear correlation function, $\xi_\pm$, the aperture-mass
dispersion, $\langle M_ {\rm ap} \rangle ^2$ \citep{1998MNRAS.296..873S}, and
the COSEBIs, $E_n$ \citep{COSEBIs}.  These statistics are linearly related to
the power spectrum via integrals of the form,
%
\begin{align}
%\label{eqn:integ}
& E_n =\int_0^{\infty}\frac{\d \ell \ell}{2\pi} W_n(\ell) P(\ell)\;,\\ \nonumber
& \langle M_ {\rm ap} \rangle ^2(\theta)=\int_0^{\infty}\frac{\d \ell \ell}{2\pi} U^2_\theta(\ell) P(\ell)\;,
\end{align}
%
where $W_n(\ell)$ and $U^2_\theta(\ell)$ are defined in \cite{COSEBIs} and
\cite{1998MNRAS.296..873S}. The corresponding equations for $\xi_\pm$ are given
in equation~\ref{eqn:xiGG}. Figure~\ref{fig:filters} shows the integrands of
these statistics for two cases, normalised to their maximum value, where the
integrands are of the form $\ell F(\ell) P(\ell)$. For $\xi_\pm$, $P(\ell)$ is
equal to the sum of the E and B-mode power spectra, motivating the development
of the aperture-mass dispersion statistic, $\langle M_ {\rm ap} \rangle ^2$.
This statistic is, however, a lossy conversion and is biased by small angular
separations, where blending of galaxies makes shear measurement challenging
\citep{KSE06}.  The COSEBIs statistic tackles both these
shortcomings.  The upper two panels of figure~\ref{fig:filters} show the
integrands\footnote{Note that the lower left panel of Fig.~1 in
\citet{2016arXiv161104954K} shows an integrated form of this function for a
maximum angular separation of $100^\prime$. However, in \citet{CFHTLenS-2pt-notomo},
the data used in \citet{2016arXiv161104954K}, $\theta$ is between $0.8'$ and
$350'$.} of $\xi_\pm$ for $\theta=100'$ and $\theta=350'$. The lower middle
panel in Figure~\ref{fig:filters} shows the COSEBIs integrands for two angular
ranges, $[1',100']$ and $[0.8',350']$, where we only show the integrands for
the lowest COSEBIs mode, $E_1$, as the higher modes generally probe larger
$\ell$-modes.  Finally, the lowest panel shows the integrands of aperture mass
dispersion statistics, for the same two maximum angular ranges.

@@

From Figure~\ref{fig:filters} we can see that the two-point cosmic shear
statistics tested by \citet{CFHTLenS-2pt-notomo} exhibit different dependences
between the angular scales sampled and the $\ell$-range probed.   If a
significant bias had been introduced at low-$\ell$ by using flat-sky and Limber
approximations, we would then expect to see a systematic shift between the
different two-point statistics with the COSEBIs statistic being essentially
unaffected as it only includes modes with $\ell \gtrsim 20$.  This is found not
to be the case with all three statistics finding $\sigma_8
(\Omega_m/0.27)^\alpha = 0.79$ with errors that range from $0.03$ to $0.06$ for
the full five-parameter fit, and $\alpha$ ranging from $0.59$ to 0.7 \citep[see
Table 5 of][]{CFHTLenS-2pt-notomo}.  This comparison further supports our
argument that the approximations highlighted by \citet{2016arXiv161104954K}
have negligible impact for current surveys.
